Number recognition for use in a Sudoku solver

### 2012-03-15 doc/icml2010stylefiles/paper.tex
fixed problem with hyperref
- finished gridSearchSVM to work with the real data
- worked on doc
- moved resource files from doc one folder up
- removed ex4.pdf

% As of 2010, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2010} with
% \usepackage[nohyperref]{icml2010} above.
\usepackage{hyperref}

@@

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
% \usepackage{icml2010} 
% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Appearing in''
\usepackage[accepted]{icml2010}

@@

\section{Introduction}
Number and letter recognition is key in many real-world applications. Digitalizing print media, automated sorting of letters, or recognition of signs in robotics or driver assistance systems require well performing algorithms for character recognition. In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated Sudoku solver.

\section{Overview}
% trim=l b r t
\begin{figure}[ht]
	\centerline{\includegraphics[width=\columnwidth, trim=25 30 25 30]{overview}}
	\caption{Overview}
	\label{fig:overview}
\end{figure}

\section{Generation of Training Data}
\subsection{Image Generation}
\begin{figure}[ht]
	\centerline{\includegraphics[scale=0.25]{examples}}
	\caption{Examples of generated numbers}
	\label{fig:generated_numbers}
\end{figure}
\subsection{Image Processing}

\section{Training}
\subsection{Neural Network}
\begin{figure}[ht]
	\centerline{\includegraphics[width=\columnwidth, trim=0 30 0 30]{nn}}
	\caption{Neural Network with one hidden layer}
	\label{fig:nn}
\end{figure}
\subsection{Support Vector Machines}
The next technique we applied to the discussed problem are Support Vector Machines \cite{Bishop200608} which try to solve the following optimization problem:
\begin{align*}
\min_{w,b} &~~ \frac{1}{2}||w||^2 + C \sum_{i=1}^{n} \xi_i\\
\text{subject to } &~~ y_i(w^T \phi(x_i) + b) - 1 + \xi_i \ge 0\\
\text{and} &~~ \xi_i \ge 0.
\end{align*}
Using a radial basis function (RBF) as kernel we performed a grid search over $C$ and $\gamma$ to find those parameters that provide the highest accuracy on the validation set. We used the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing the grid search.
\subsection{Logistic Regression}

\section{Image Recognition and Preprocessing}

\section{Number Prediction}

\section{Evaluation}

\section{Future Work}
\color{red} include information if a sudoku is unsolvable \color{black}

>>>>

% As of 2010, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2010} with
% \usepackage[nohyperref]{icml2010} above.
%\usepackage{hyperref}

@@

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
% \usepackage{icml2010} 
% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Appearing in''
\usepackage[accepted, nohyperref]{icml2010}

@@

\section{Introduction}
Number and letter recognition is key in many real-world applications. Digitalizing print media, automated sorting of letters, or recognition of signs in robotics or driver assistance systems require well performing algorithms for character recognition. In the following we analyze the performance of different techniques for number recognition, namely Neural Networks (NN), Support Vector Machines (SVM) and Logistic Regression (LR), by using them in an automated Sudoku solver.

\section{Overview}
Figure \ref{fig:overview} shows the basic data flow and the participating data manipulators. First of all we generate data by sampling position, rotation, size and color of a number and place it based on these values on a white 20x20 pixel image. Examples of 9 such generated images are shown in Figure \ref{fig:generated_numbers}. These images are preprocessed by an octave script and written to a matrix. The corresponding labels are defined by the name of the folder the images are stored in.
% trim=l b r t
\begin{figure}[ht]
	\centerline{\includegraphics[width=\columnwidth, trim=25 30 25 30]{../overview}}
	\caption{Overview}
	\label{fig:overview}
\end{figure}
The $X^{n\text{x}400}$ and the target vector $y$ are the inputs for training NN, SVM and LR, as described in Section \ref{training}. Trained models/parameters are saved in ``.mat'' files that can be loaded by the number prediction. The other input for number prediction is an input matrix of the same form as $X$ with a row entry for each Sudoku field that was detected to contain a number.

\section{Generation of Training Data}
\subsection{Image Generation}
\begin{figure}[ht]
	\centerline{\includegraphics[scale=0.25]{../examples}}
	\caption{Examples of generated numbers}
	\label{fig:generated_numbers}
\end{figure}
\subsection{Image Processing}

\section{Training}
\label{training}
\subsection{Neural Network}
Applying a NN seems promising as it was already successfully applied to the quite similar problem of detecting handwritten digits. \color{red}ref!\color{black} For the NN we choose a simple architecture with 25 hidden sigmoidal units (+ bias unit), as shown in figure \ref{fig:nn}. The size of the input layer is 400 (given by the 20 x 20 pixel images) and the size of the output layer is defined by the number of classes, in this case the numbers 1 to 9. This simple architecture turned out to perform pretty well compared to more complex ones with a higher number of hidden layers. Training is performed with error backpropagation, thus minimizing the error function
\begin{align*}
& E(w) = \sum_{n=1}^N E_n(w)\\
\text{with}~~E&_n(w) = \frac{1}{2}\sum_k (y_{nk}-t_{nk})\\
\text{and}~~~~~&~~~y_k = \sum_i w_{ki}x_{i}\\
\end{align*}
subject to $w$.
\begin{figure}[ht]
	\centerline{\includegraphics[width=\columnwidth, trim=0 30 0 30]{../nn}}
	\caption{Neural Network with one hidden layer}
	\label{fig:nn}
\end{figure}
For NNs we test two implementations. On the one hand we use a partly self-developed octave code for two-layer NNs which was written as part of the stanford machine learning course\footnote{http://www.ml-class.org}, on the other hand we include PyBrain \cite{schaul2010pybrain} for comparison and easier evaluation of more sophisticated network architectures. To find a good value for the regularization parameter $\lambda$ we perform a grid search with k-fold cross-validation which gives us a good performance when choosing $\lambda = 1$.

\subsection{Support Vector Machines}
The next technique we apply to the discussed problem are Support Vector Machines \cite{Bishop200608} which try to solve the following optimization problem:
\begin{align*}
\min_{w,b} &~~ \frac{1}{2}||w||^2 + C \sum_{i=1}^{n} \xi_i\\
\text{subject to } &~~ y_i(w^T \phi(x_i) + b) - 1 + \xi_i \ge 0\\
\text{and} &~~ \xi_i \ge 0.
\end{align*}
Using a radial basis function (RBF) as kernel we perform a grid search over $C$ and $\gamma$ to find those parameters that provide highest accuracy on the test set. We use the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing grid search with k-fold cross-validation. Grid search with cross validation resulted in accuracies illustrated in Figure \ref{fig:svm_grid_search}. From the best parameter pairs we choose the one that performs best on the validation set, which gives us \color{red}$C = ?$ and $\lambda = ?$.\color{black}
\begin{figure}[ht]
	\centerline{\includegraphics[scale=0.20, trim=0 0 10 0]{../gridSearchSVMPlot}}
	\caption{}
	\label{fig:svm_grid_search}
\end{figure}

\subsection{Logistic Regression}
Finally we apply Logistic Regression to the datasets. Again we use a (vectorized) octave implementation for learning and prediction. Therefore we minimize the (regularized) error function
\begin{align*}
E(w) &= \frac{1}{n}(-y^T ln(\Phi w) + (1-y^T)ln(1-\Phi w))\\
&~~~+\frac{\lambda}{2n}w^Tw
\end{align*}
with respect to $w$. Minimization is performed by an algorithm\footnote{http://sprinkler.googlecode.com/svn/trunk/\\regression/fmincg.m} that was provided during the stanford machine learning course. Grid search and k-fold cross-validation recommend a value for the regularization parameter $\lambda$.

\section{Image Recognition and Preprocessing}
Image recognition is written in C by extensive use of the open source library OpenCV\footnote{http://opencv.willowgarage.com/wiki/}. Images captured by the webcam are continously analyzed to find a Sudoku square. Images are transformed to greyscale, then an adaptive threshold algorithm is used to preprocess the image for contour detection by OpenCV. OpenCV returns a list of contours which are further checked for certain criteria like having four corners or being of certain size. Since the contours are not detected precise enough lines are fitted through six checkpoints on each edge in order to achieve subpixel accuracy. The content enclosed by the corners is perspectively warped to fit a predefined square size. Finally, the thresholded Sudoku field is saved to disk. A python script is then used to split the whole field into 9x9 subfields with each subfield being flood filled around the corners for noise reduction. The format of the subfields now equals to the one provided by the generated images which allows us to apply the trained NN/SVM/LR models and, finally, apply a Sudoku solver.

\section{Evaluation}
For evaluation of the different techniques we define the following datasets:
\begin{itemize}
\item d1: 4500 generated images
\item d2: $\sim$ 1000 gathered images
\item d3: D1 $\cup$ D2
\item dVal: images of a captured Sudoku
\end{itemize}

\color{red}
\begin{table}[t]
\caption{Classification accuracies for Neural Network, SVM and Logistic Regression on different data sets.}
\label{sample-table}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{lcccc}
\hline
\abovespace\belowspace
Data set & NN & SVM & LR & Winner \\
\hline
\abovespace
D1		& 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& 96.7$\pm$ 0.2& ?\\
D2		& 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& ? & ?\\
D3		& 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& ? & ? \\
\belowspace
DVal	& 74.8$\pm$ 0.5& 78.3$\pm$ 0.6& ? & ?        \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}
\color{black}

\section{Future Work}
\color{red}
\begin{itemize}
\item include information if a sudoku is unsolvable
\item online learning
\end{itemize}
\color{black}
